---
title: "Data 502 Midterm#2"
author: "Travis McKenzie"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(tidyverse)
library(maps)
library(tigris)
options(tigris_use_cache = TRUE)
library(colorspace)
library(plotly)

```

### Midterm 2 - 11/15/23
Read in the data


```{r county & k-12}

county<-read.csv("https://raw.githubusercontent.com/kitadasmalley/Teaching/main/DATA502/FA2023/Data/immuneOrCounty.csv")
k12 <-read.csv("https://raw.githubusercontent.com/kitadasmalley/Teaching/main/DATA502/FA2023/Data/immune_rates.csv")
k12$Per_Vac_AllVac<-as.numeric(k12$Per_Vac_AllVac)


```

## Review Data


```{r summary, echo=FALSE}
head(county)
head(k12)
unique(county$County)

```

##Front End
Load Maps Package and Map Data, Join Data, Create new vectors


```{r}
or_counties = map_data("county", "oregon") %>% 
  select(lon = long, lat, group, id = subregion)

county = county %>%
  mutate(id=tolower(County))

ORjoin=full_join(county, or_counties)

ORjoin=ORjoin%>%
  mutate(all_pct_cut=cut(all_pct, breaks = c(97.1,94.39,92.3,88.6,85)))%>%
  mutate(all_pct_cut1=cut(all_pct, breaks = c(0,90,100)))

k12=k12%>%
  mutate(cutVac=cut(Per_Vac_AllVac, breaks=c(100,95,90,80,10)))%>%
  mutate(cutVac1=cut(Per_Vac_AllVac, breaks=c(0,90,100)))

```
## Construct Graph
Basics and experimentation

```{r}
ORjoin%>%
ggplot()+
  geom_polygon(aes(lon, lat, group=group, fill=all_pct))+
  coord_quickmap()
```


```{r}
ORjoin%>%
ggplot()+
  geom_polygon(aes(lon, lat, group=group, fill=all_pct))+
  coord_quickmap()+
  geom_point(data=k12, aes(X,Y))
  

```

```{r}
ORjoin%>%
ggplot()+
  geom_polygon(aes(lon, lat, group=group, fill=all_pct))+
  coord_quickmap()+
  geom_point(data=k12, aes(X,Y, color=Per_Vac_AllVac))+
  scale_fill_continuous_sequential("Teal", limits=c(85,100))

```

#Final Recreation:

```{r}
ORjoin%>%
ggplot()+
  geom_polygon(color="darkgray",aes(lon, lat, group=group, fill=all_pct_cut))+
  coord_quickmap()+
  geom_point(data=k12, aes(X,Y, color=cutVac), na.rm=FALSE)+
  scale_fill_discrete_sequential("Teal", labels=c("85-88.6", ">88.6-92.3", ">92.3-94.39", ">94.39-97"), name="County Immunization Rates
All Vaccines")+
scale_color_discrete_sequential("Heat", na.value="darkgray", order=c(1:4), 
                                labels=c("Safest (95-100% of students fully
vaccinated)", 
                                         "Moderately Vulnerable (90-94.9% 
of students fully vaccinated)", 
                                         "More Vulnerable (80-89.9% of students
fully vaccinated)", 
                                         "Most Vulnerable(less than 80.0% of 
students fully vaccinated)", 
                                         "Site has fewer than 10 children. 
No data available"), 
name="K12 Schools
% Vaccinated (All Vaccines)")+
  labs(title="   How well-vaccinated is your child care, school and county?",
       caption="Data from Oregon.gov")+
  theme_void()
```
##Interactivity Attempt

```{r}
p=ORjoin%>%
ggplot(aes(County=County))+
  geom_polygon(color="darkgray",aes(lon, lat, group=group, fill=all_pct_cut))+
  coord_quickmap()+
  geom_point(data=k12, aes(X,Y, color=cutVac, 
                 text=paste(
                            "Facility:", Site.Name, "<br>",
                            "District:", School_District, "<br>",
                            "Vaccine %", Per_Vac_AllVac)))+
  scale_fill_discrete_sequential("Teal",guide="none")+
scale_color_discrete_sequential("Heat", na.value="darkgray", order=c(1:4), guide="none")+
  labs(title="   How well-vaccinated is your child care, school and county?",
       caption="Data from Oregon.gov")+
  theme_void()+
  theme(
    legend.position = "none"
  )

ggplotly(p, tooltip=c("County","text"))


```

#Alternative Creation

```{r}
ORjoin%>%
ggplot()+
  geom_polygon(color="lightgray", linewidth=1.2,aes(lon, lat, group=group, fill=all_pct_cut1))+
  coord_quickmap()+
  geom_point(data=k12, aes(X,Y, color=cutVac1))+
  scale_fill_discrete_sequential("Light Grays",order=c(2,1), 
      labels=c("Sub-Optimal Vaccination Levels
(below 90%)", "Optimal Vaccination Levels
(at or above 90%)"),
name="County Immunization Rates
(All Vaccines)")+
  scale_color_discrete_sequential("Terrain",na.value="gray", order=c(2,1), 
      labels=c("Sub-Optimal Vaccination Levels
(below 90%)",
               "Optimal Vacciation Levels
(at or above 90%)",
               "Insufficient Data
(below 10%)"),
name="K12 Schools
% Vaccinated
(All Vaccines)")+
  labs(title="Five Oregon Counties Need Vaccine Support",
       subtitle = "State department recommends providing support to these counties to enhance immunization rates", caption="Data from Oregon.gov")+
  theme_void()

```

###Conclusion
For my alternative, I wanted to tell a story about establishing vaccine availability, education and support to areas that are most likely to be in need.  I established a threshold of 90% as my differentiation between optimal and sub-optimal.  I tested a litany of colors on this map many times and am partially satisfied with the result.  Would also like to figure out how I can draw borders around certain counties to better highlight them and lean on enclosure.  I felt like using a gray palette against a contrasting hue like green allowed me to highlight particular counties to address a concern.



